PANCHANGE_data_preprocessing

### 2020-08-05 PANCHANGE_data_preprocessing2020-07-15.Rmd
creating COVID respiratory symptom binary categories

```{r distribution of COVID symptoms}
#persistent cough
coping.glad.respiratory.raw.numeric %>%
  ggplot(
    mapping = aes(
      y = respiratory.persistent_cough_numeric 
    )
  ) +
  geom_bar() +
  labs(
    title = "Persistant cough"
  )

@@

Choosing cut-offs for the likert scale symptoms of "no symptoms", "mild symptoms", "moderate symptoms" and "severe symptoms" for the fatigue, persistent cough and skipped meals variables. 
Fatigue: "Severe symptoms" as cut off
Persistent cough: "Mild symptoms" as cut off
Skipped meals: "Moderate symptoms" as cut off
```{r proportion of people who report COVID symptoms in Menni et al}
#fatigue
percentage_fatigue <- ((0.298*6452) + (0.1556*9186) + (0.0871*2434931))/(6452+9186+2434931) 
percentage_fatigue   # 0.08791204% of Menni et al(2020) have fatigue
quantile(x = coping.glad.respiratory.raw.numeric$respiratory.feeling_tiredfatigued_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_fatigue) #this gives 8% of the sample reporting fatigue when using the "severe symptoms" fatigue response

#persistent cough
percentage_pcough <- ((0.5673*6452) + (0.4556*9186) + (0.3017*2434931))/(6452+9186+2434931)
percentage_pcough    # 0.3029762% of Menni et al(2020) have fatigue
quantile(x = coping.glad.respiratory.raw.numeric$respiratory.persistent_cough_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_pcough) # this gives 0.3029762% of the sample reporting fatigue when using the "mild symptoms" persistent cough response

@@

>>>>

#Distribution of likert scale COVID-19 symptoms in COPING
```{r distribution of COVID symptoms}
#persistent cough
coping.glad.respiratory.raw.numeric %>%
  ggplot(
    mapping = aes(
      y = respiratory.persistent_cough_numeric 
    )
  ) +
  geom_bar() +
  labs(
    title = "Persistant cough"
  )

@@

#Choosing cut-offs for the likert scale symptoms of "no symptoms", "mild symptoms", "moderate symptoms" and "severe symptoms" for the fatigue, persistent cough and skipped meals variables. 
Coding:
0 = No symptoms
1 = Mild symptoms
2 = Moderate symptoms
3 = Severe symptoms

Results:
Persistent cough: "Mild symptoms" as cut off
Fatigue: "Severe symptoms" as cut off
Skipped meals: "Moderate symptoms" as cut off
```{r proportion of people who report COVID symptoms in Menni et al}
#persistent cough
percentage_pcough <- ((0.5673*6452) + (0.4556*9186) + (0.3017*2434931))/(6452+9186+2434931)
percentage_pcough    # 0.3029762% of Menni et al(2020) have fatigue
quantile(x = coping.glad.respiratory.raw.numeric$respiratory.persistent_cough_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_pcough) # this gives 0.3029762% of the sample reporting fatigue when using the "mild symptoms" persistent cough response

#fatigue
percentage_fatigue <- ((0.298*6452) + (0.1556*9186) + (0.0871*2434931))/(6452+9186+2434931) 
percentage_fatigue   # 0.08791204% of Menni et al(2020) have fatigue
quantile(x = coping.glad.respiratory.raw.numeric$respiratory.feeling_tiredfatigued_numeric, 
         na.rm = TRUE, 
         probs = 1-percentage_fatigue) #this gives 8% of the sample reporting fatigue when using the "severe symptoms" fatigue response

@@

Creating binary groups for COVID symptoms
```{r creating binary groups for COVID symptoms}
#loss or smell or taste
coping.glad.respiratory.raw.numeric <- coping.glad.respiratory.raw.numeric %>%
  mutate(
    respiratory.anosmia_binary =
      if_else(
        respiratory.was_your_sense_of_smell_impaired_numeric == 1 | respiratory.was_your_sense_of_taste_impaired_numeric == 1, 
        1,
        0
      )
  )

freq(coping.glad.respiratory.raw.numeric$respiratory.anosmia.binary)

#persistent cough
coping.glad.respiratory.raw.numeric <- coping.glad.respiratory.raw.numeric %>%
  mutate(
    respiratory.pcough_binary =
      if_else(
        respiratory.persistent_cough_numeric >= 1, #those who score 1 or above (mild or above)
        1,
        0
      )
  )

#fatigue
coping.glad.respiratory.raw.numeric <- coping.glad.respiratory.raw.numeric %>%
  mutate(
    respiratory.fatigue_binary =
      if_else(
        respiratory.feeling_tiredfatigued_numeric >= 3, #those who score 3 (severe symptoms)
        1,
        0
      )
  )

#skipped meals
coping.glad.respiratory.raw.numeric <- coping.glad.respiratory.raw.numeric %>%
  mutate(
    respiratory.skipped_meals_binary =
      if_else(
        respiratory.skipping_meals_because_you_felt_unwell_numeric >= 2,
        1,
        0
      )
  )
```
